Do we need detailed demographic data to forecast plant population responses to climate change?

### 2015-06-15 manuscript/manuscript.Rmd
updates ms with new cliamte change figure

All R code and data necessary to replicate our analysis is archived on GitHub as release 1.0 (http://github.com/atredennick/MicroMesoForecast/releases). That stable release will remain static as a record of this analysis, but subsequent versions may appear if we update this work.

@@

##  QBM results
# Read in data
qbm_onestep <- readRDS("../analysis/quadBM/simulations/validation_results/qbm_one-step_forecasts_combined.rds")
# Average predictions over quad-year reps
avg_qbm <- ddply(qbm_onestep, .(species, quad, year), summarise,
                 avg_prediction = median(cover.t1),
                 observation = mean(cover.t0),
                 quant_dist = quantile(cover.t1*100, probs = 0.95) - quantile(cover.t1, probs = 0.05))
# Calculate error
avg_qbm$error <- avg_qbm$observation*100 - avg_qbm$avg_prediction*100
# Aggregate over species
qbmstats <- ddply(avg_qbm, .(species), summarise,
                  mae = mean(abs(error)),
                  mean_cover = mean(observation*100),
                  quant_dist = mean(quant_dist))
qbmstats$model <- "QBM"

@@

colnames(stats_table) <- c("Species", "Model", "MAE", "90% Distance", "Mean Obs. Cover")
print.xtable(xtable(stats_table, caption = "Mean absolute error (MAE) and accuracy (Pearson's $\\rho$) for one-step-ahead forecast from both model types. Forecasts were made without random year effects; only climate covariates could explain year-to-year variation. 90 Distance refers to the average distance between the upper and lower 90 percentiles of the 100 predicted values for each quadrat-year combination."),type="latex", comment=FALSE,
             include.rownames=FALSE, caption.placement="top")
```
The IPM MAE is significantly lower for HECO ($p = 3.3 \times 10^{-10}$) and POSE ($p = 0.00013$). MAEs are statisticially similar between models for BOGR and PASM.

@@

Figures
-------
```{r figure_1,dependson="plot-options", fig.cap="Work flow of the data aggregation, model fitting, and population simulating.", fig.height=6}
library(png)
library(grid)
img <- readPNG("components/figure/micromeso_flow.png")
grid.raster(img)
```

@@

```{r figure_3, dependson="plot-options", fig.cap="Proportional change in species' mean cover caused by a 1% increase in observed precipitation, temperature, or both as predicted by the individual-based IPM and the aggregate-based QBM. Note that the change for POSE due to a precipitation increase predicted by the QBM is almost zero and so does not show up in the figure. Labels on x-axis refer to: 'ppt' = 1% increase in mean precipitation; 'temp' = 1% increase in mean temperature; 'ppt + temp' = 1% increase in mean precipitation and mean temperature.", fig.width=8.5, fig.height=3, cache=FALSE, results='hide'}
# ipm_diffs <- readRDS("../montana/ipm/simulations/results/ipm_climatesims_logdiffs.RDS")
# ipm_diffs$model <- rep("IPM", nrow(ipm_diffs))
# ipm_diffs$type <- rep("paramunc", nrow(ipm_diffs))
# ipm_diffs2 <- readRDS("../montana/ipm/simulations/results/ipm_climatesims_logdiffs_mean.RDS")
# ipm_diffs2$model <- rep("IPM", nrow(ipm_diffs))
# ipm_diffs2$type <- rep("parammean", nrow(ipm_diffs))
# ipm_diffs$med_cover <- ipm_diffs2$med_cover
# ipm_diffs <- rbind(ipm_diffs, ipm_diffs2)

# ggplot(ipm_diffs)+
#   geom_hline(aes(yintercept=0), linetype=2)+
#   geom_errorbar(aes(x=species, ymin=lo_cover, ymax=up_cover, group=variable),
#                 position=dgd, width=0.25, color="grey45")+
#   geom_point(aes(x=species, y=med_cover, shape=variable),
#              position=dgd, size=5, color="white")+
#   geom_point(aes(x=species, y=med_cover, shape=variable),
#              position=dgd, size=3)+
# #   scale_shape_manual(values = c(19,17), name="Model type")+
#   xlab("Climatic change")+
#   ylab("Proportional change in percent cover")+
#   facet_wrap("type",scales = "free")+
# #   scale_x_discrete(labels=c("ppt", "temp", "ppt + tmp"))+
#   theme_bw()

# qbm_diffs <- readRDS("../montana/quadBM/simulations/results/qbm_climatesims_logdiffs.RDS")
# qbm_diffs$type <- rep("QBM", nrow(qbm_diffs))
# ipm_diffs <- readRDS("../montana/ipm/simulations/results/ipm_climatesims_logdiffs.RDS")
# ipm_diffs$type <- rep("IPM", nrow(ipm_diffs))
# 
# diff_df <- rbind(qbm_diffs, ipm_diffs)
# 
# myCols2 <- c("grey45", "#277BA8", "#7ABBBD", "#AED77A")
# dgd = position_dodge(width = 0.60)
# ggplot(diff_df)+
#   geom_hline(aes(yintercept=0), linetype=2)+
#   geom_errorbar(aes(x=variable, ymin=lo_cover, ymax=up_cover, group=type),
#                 position=dgd, width=0.25, color="grey45")+
#   geom_point(aes(x=variable, y=med_cover, shape=type),
#              position=dgd, size=5, color="white")+
#   geom_point(aes(x=variable, y=med_cover, shape=type),
#              position=dgd, size=3)+
#   scale_shape_manual(values = c(19,17), name="Model type")+
#   xlab("Climatic change")+
#   ylab("Proportional change in percent cover")+
#   facet_grid(.~species)+
#   scale_x_discrete(labels=c("ppt", "temp", "ppt + tmp"))+
#   theme_bw()
```

.

>>>>

All R code and data necessary to reproduce our analysis is archived on GitHub as release 1.0 (http://github.com/atredennick/MicroMesoForecast/releases). That stable release will remain static as a record of this analysis, but subsequent versions may appear if we update this work.

@@

##  QBM results
# Read in data
qbm_onestep <- readRDS("../analysis/quadBM/simulations/results/one_step_validation/qbm_one-step_forecasts_combined.rds")
# Average predictions over quad-year reps
avg_qbm <- ddply(qbm_onestep, .(species, quad, year), summarise,
                 avg_prediction = median(cover.t1),
                 observation = mean(cover.t0),
                 quant_dist = quantile(cover.t1*100, probs = 0.95) - quantile(cover.t1, probs = 0.05))
# Calculate error
avg_qbm$error <- avg_qbm$observation*100 - avg_qbm$avg_prediction*100
# Aggregate over species
qbmstats <- ddply(avg_qbm, .(species), summarise,
                  mae = mean(abs(error)),
                  mean_cover = mean(observation*100),
                  quant_dist = mean(quant_dist))
qbmstats$model <- "QBM"

@@

colnames(stats_table) <- c("Species", "Model", "MAE", "90% Distance", "Mean Obs. Cover")
print.xtable(xtable(stats_table, caption = "Mean absolute error (MAE) and accuracy (Pearson's $\\rho$) for one-step-ahead forecast from both model types. Forecasts were made without random year effects; only climate covariates could explain year-to-year variation. 90\\% Distance refers to the average distance between the upper and lower 90th percentiles of the 100 predicted values for each quadrat-year combination."),type="latex", comment=FALSE,
             include.rownames=FALSE, caption.placement="top")
```
The IPM MAE is significantly lower for HECO ($p = 3.3 \times 10^{-10}$) and POSE ($p = 0.00013$). MAEs are statisticially similar between models for BOGR and PASM.

@@

Figures
-------
```{r figure_1,dependson="plot-options", fig.cap="Work flow of the data aggregation, model fitting, and population simulating.", fig.height=6}
library(png)
library(grid)
img <- readPNG("components/figure/micromeso_flowchart.png")
grid.raster(img)
```

@@

```{r figure_3, dependson="plot-options", fig.cap="Proportional change in species' mean cover caused by a 1% increase in observed precipitation (+ppt), temperature (+temp), or both (+ppt&temp) as predicted by the individual-based IPM and the aggregate-based QBM. Top panels show the mean predicted proportional change in cover; lower panels show the 90% quantiles of predicted proportional cover change. Mean predictions and their associated uncertainties are shown in separate panels for visual clarity.", fig.width=8.5, fig.height=6, cache=FALSE, results='hide'}
setwd("~/Repos/MicroMesoForecast/analysis/")

####
##  Mean parameter runs
####
meanparam_ipm <- data.frame(cover=NA, species=NA, 
                             climsim=NA, model=NA)
##  IPM 
path <- "./ipm/simulations/results/climchange_meanparams/"
ipmfiles_mean <- list.files(path)
ipmfiles_mean <- ipmfiles_mean[grep("cover", ipmfiles_mean)]
for(file in ipmfiles_mean){
  tmp <- read.csv(paste(path,file,sep=""))
  tmp$model <- "IPM"
  meanparam_ipm <- rbind(meanparam_ipm, tmp[,c("cover", "species", "climsim", "model")])
}
meanparam_ipm <- meanparam_ipm[2:nrow(meanparam_ipm),]

newd <- data.frame(cover=NA, species=NA, 
                   climsim=NA, model=NA, obscover=NA)
for(dospp in unique(meanparam_ipm$species)){
  tmp <- subset(meanparam_ipm, species==dospp)
  obscover <- subset(tmp, climsim=="obs")["cover"]
  tmp2 <- subset(tmp, climsim!="obs")
  tmp2$obscover <- rep(obscover$cover, times = length(unique(tmp2$climsim)))
  newd <- rbind(newd, tmp2)
}
ipm_meansims <- newd[2:nrow(newd),]
ipm_meansims$propdiff <- with(ipm_meansims, log(cover*100)-log(obscover*100))

##  QBM
meanparam_qbm <- data.frame(cover=NA, species=NA, 
                             climsim=NA, model=NA)
path <- "./quadBM/simulations/results/climatechange_meanparams/"
qbmfiles_mean <- list.files(path)
for(file in qbmfiles_mean){
  tmp <- readRDS(paste(path,file,sep=""))
  tmp$model <- "QBM"
  meanparam_qbm <- rbind(meanparam_qbm, tmp[,c("cover", "species", "climsim", "model")])
}
meanparam_qbm <- meanparam_qbm[2:nrow(meanparam_qbm),]

newd <- data.frame(cover=NA, species=NA, 
                   climsim=NA, model=NA, obscover=NA)
for(dospp in unique(meanparam_qbm$species)){
  tmp <- subset(meanparam_qbm, species==dospp)
  obscover <- subset(tmp, climsim=="obs")["cover"]
  tmp2 <- subset(tmp, climsim!="obs")
  tmp2$obscover <- rep(obscover$cover, times = length(unique(tmp2$climsim)))
  newd <- rbind(newd, tmp2)
}
qbm_meansims <- newd[2:nrow(newd),]
qbm_meansims$propdiff <- with(qbm_meansims, log(cover*100)-log(obscover*100))

##  Merge IPM and QBM mean parameter results
bothmodel_outs <- rbind(ipm_meansims, qbm_meansims)
meanparam_agg <- ddply(bothmodel_outs, .(species, climsim, model), summarise,
                       meandiff = mean(propdiff),
                       hidiff = quantile(propdiff, probs = 0.95),
                       lodiff = quantile(propdiff, probs = 0.05))
dgd = position_dodge(width = 0.9)
meanplot <- ggplot(meanparam_agg)+
  geom_bar(aes(x=climsim, y=meandiff, fill=model),
             position=dgd, stat="identity")+
  theme_bw()+
  scale_fill_manual(values=c("black", "grey45"), name="Model")+
  facet_wrap("species", scales="free", nrow=1)+
  scale_x_discrete(labels=c("+ppt", "+temp", "+ppt&temp"))+
  ylab("Mean proportional difference")+
  xlab("Climate simulation")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

dgd = position_dodge(width = 0.6)
uncertainplot <- ggplot(meanparam_agg)+
  geom_errorbar(aes(x=climsim, ymax=hidiff, ymin=lodiff, group=model, linetype=model),
                position=dgd, width=0.25)+
  theme_bw()+
  scale_linetype_discrete(name="Model")+
  facet_wrap("species", nrow=1, scales="free")+
  scale_x_discrete(labels=c("+ppt", "+temp", "+ppt&temp"))+
  ylab("Proportional difference")+
  xlab("Climate simulation")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

meanparamplot <- grid.arrange(meanplot, uncertainplot, ncol=1, nrow=2)
print(meanparamplot)
```

### 2015-06-15 manuscript/manuscript.tex
updates ms with new cliamte change figure

\documentclass[12pt,]{article}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
    \setmainfont{times}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
% Redefine \includegraphics so that, unless explicit options are
% given, the image width will not exceed the width of the page.
% Images get their normal width if they fit onto the page, but
% are scaled down if they would overflow the margins.
\makeatletter
\def\ScaleIfNeeded{%
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother
\let\Oldincludegraphics\includegraphics
{%
 \catcode`\@=11\relax%
 \gdef\includegraphics{\@ifnextchar[{\Oldincludegraphics}{\Oldincludegraphics[width=\ScaleIfNeeded]}}%
}%
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{0}

\author{}
\date{}
\usepackage{lineno}
\linenumbers
\usepackage{setspace}
\doublespacing
\setlength\parindent{24pt}

@@

\subsubsection[Andrew T. Tredennick and Peter B. Adler]{Andrew T.
Tredennick\footnote{Corresponding author:
  \href{mailto:atredenn@gmail.com}{\href{mailto:atredenn@gmail.com}{atredenn@gmail.com}}}
and Peter B.
Adler}\label{andrew-t.-tredennickcorrauth-and-peter-b.-adler}

@@

Our demographic data comes from the Fort Keogh Livestock and Range
Research Laboratory in eastern Montana's northern mixed prairie near
Miles City, Montana, USA ($46^{\circ}$ 19' N, $105^{\circ}$ 48' W). The
dataset is freely available on Ecological Archives\footnote{\url{http://esapubs.org/archive/ecol/E092/143/}}
(Anderson et al. 2011) , and interested readers should refer to the
metadata therein for a complete description. The site is about 800 m
above sea level and mean annual precipitation (1878-2009) is 334 mm,
with most annual precipitation falling from April through September. The
site is grass dominated and, for the purposes of our study, we focus on
the four most abundant graminoid species: \emph{Bouteloua gracilis}
(BOGR), \emph{Hesperostipa comata} (HECO), \emph{Pascopyrum smithii}
(PASM), and \emph{Poa secunda} (POSE).

From 1932 to 1945 individual plants were identified and mapped annually
in 44 1-$\text{m}^2$ quadrats using a pantograph. The quadrats were
distributed in six pastures, each assigned a grazing treatment of light
(1.24 ha/animal unit month), moderate (0.92 ha/aum), and heavy (0.76
ha/aum) stocking rates (two pastures per treatment). In this analysis we
account for potential differences among the grazing treatments, but do
not focus on grazing$\times$climate interactions. The annual maps of the
quadrats were digitized and the fates of individual plants tracked and
extracted using a computer program. Daily climate data, which we
aggregated into climate variables of interest, are available for the
duration of the data collection period (1932 - 1945) from the Miles City
airport, Wiley Field, 9 km from the study site.

@@

At both levels of inference (individual and quadrat), the building
blocks of our population models are vital rate regressions. For
individual level data we fit models for survival, growth, and
recruitment of new individuals for each species. At the quadrat level we
fit a single regression model for population growth. We describe the
statistical models separately since fitting the models required
different approaches. All models contain five climate covariate that we
chose \emph{a priori}: ``water year'' precipitation at \emph{t}-1
(lagppt); fall through spring precipitation at \emph{t}-1 and \emph{t}-2
(ppt1 and ppt2, respectively) and mean spring temperature at \emph{t}-1
and \emph{t}-2 (TmeanSpr1 and TmeanSpr2, respectively), where \emph{t}
is the observation year. We also include interactions among same-year
climate covariates (e.g., ppt1 $\times$ TmeansSpr1) and climate $\times$
size interactions.

We fit all models using a hierarchical Bayesian approach, which we
describe in more detail below. However, for each vital rate statistical
model we also define the likelihood model we use. For the likelihood
models, \textbf{Y} is always the relevant vector of observations (e.g.,
whether a genet survived {[}1{]} or not {[}0{]} from year $t$ to $t+1$).

@@

where $x_{jq,t}$ is the log of species' $j$ proportional cover in
quadrat $q$ at time $t$ and all other parameters are as in the
individual-level growth model (Eq. \#). The log normal likelihood
includes a truncation (T{[}0,1{]}) to ensure that predicted values do
not exceed 100\% cover.

@@

We use an environmentally stochastic IPM (Rees and Ellner 2009) that
includes the random year effects and the climate covariates from the
vital rate statistical models. But note that we can, and do for some
simulations, ignore the random year effects so that only the climate
effects can drive interannual variation. Our IPM follows the
specification of Chu and Adler (2015) where the population of species
\emph{j} is a density function $n(u_{j},t)$ giving the density of
sized-\emph{u} genets at time \emph{t}. Genet size is on the natural log
scale, so that $n(u_{j},t)du$ is the number of genets whose area (on the
arithmetic scale) is between $e^{u_{j}}$ and $e^{u_{j}+du}$. So, the
density function for any size \emph{v} at time $t+1$ is

@@

where $k_{j}(v_{j},u_{j},\bar{\bold{w_{j}}})$ is the population kernal
that describes all possible transitions from size $u$ to $v$ and
$\bar{\bold{w_{j}}}$ is a vector of estimates of average crowding
experienced from all other species by a genet of size $u_j$ and species
$j$. The integral is evaluated over all possible sizes between
predefined lower (\emph{L}) and upper (\emph{U}) size limits that extend
beyond the range of observed genet sizes.

@@

which, said plainly, means we are calculating growth (\emph{G}) for
individuals that survive (\emph{S}) from time \emph{t} to \emph{t+1} and
adding in newly recruited (\emph{R}) individuals of an average sized
one-year-old genet for the focal species. Our stastical model for
recruitment (\emph{R}, described above) returns the number of new
recruit produced per quadrat. Following previous work (Adler et al.
2012, Chu and Adler 2015), we assume that fecundity increases linearly
with size
($R_j(v_{j},u_{j},\bar{\bold{w_{j}}}) = e^{u_j}R_j(v_{j},\bar{\bold{w_{j}}})$)
to incorporate the recruitment function in the spatially-implicit IPM.

@@

\begin{table}[ht]
\centering
\caption{Mean absolute error (MAE) and accuracy (Pearson's $\rho$) for one-step-ahead forecast from both model types. Forecasts were made without random year effects; only climate covariates could explain year-to-year variation. 90 Distance refers to the average distance between the upper and lower 90 percentiles of the 100 predicted values for each quadrat-year combination.} 
\begin{tabular}{llrrr}
  \hline
Species & Model & MAE & 90\% Distance & Mean Obs. Cover \\ 
  \hline
BOGR & IPM & 12.18 & 38.52 & 9.43 \\ 
  BOGR & QBM & 19.66 & 56.50 & 9.26 \\ 
  HECO & IPM & 1.22 & 6.47 & 1.15 \\ 
  HECO & QBM & 12.35 & 41.11 & 1.18 \\ 
  PASM & IPM & 0.19 & 1.65 & 0.42 \\ 
  PASM & QBM & 0.55 & 7.78 & 0.42 \\ 
  POSE & IPM & 1.37 & 7.64 & 1.25 \\ 
  POSE & QBM & 1.79 & 40.59 & 1.27 \\ 
   \hline
\end{tabular}
\end{table}

The IPM MAE is significantly lower for HECO ($p = 3.3 \times 10^{-10}$)
and POSE ($p = 0.00013$). MAEs are statisticially similar between models
for BOGR and PASM.

@@

@@

.

\subsection{References}\label{references}

>>>>

\documentclass[12pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
    \setmainfont{times}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{0}

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}
\setlength{\droptitle}{-2em}
  \title{}
  \pretitle{\vspace{\droptitle}}
  \posttitle{}
  \author{}
  \preauthor{}\postauthor{}
  \date{}
  \predate{}\postdate{}

\usepackage{lineno}
\linenumbers
\usepackage{setspace}
\doublespacing

@@

\subsubsection[Andrew T. Tredennick and Peter B. Adler]{Andrew T.
Tredennick\footnote{Corresponding author:
  \href{mailto:atredenn@gmail.com}{\href{mailto:atredenn@gmail.com}{\nolinkurl{atredenn@gmail.com}}}}
and Peter B.
Adler}\label{andrew-t.-tredennickcorrauth-and-peter-b.-adler}

@@

Our demographic data comes from the Fort Keogh Livestock and Range
Research Laboratory in eastern Montana's northern mixed prairie near
Miles City, Montana, USA (\(46^{\circ}\) 19' N, \(105^{\circ}\) 48' W).
The dataset is freely available on Ecological Archives\footnote{\url{http://esapubs.org/archive/ecol/E092/143/}}
(Anderson et al. 2011) , and interested readers should refer to the
metadata therein for a complete description. The site is about 800 m
above sea level and mean annual precipitation (1878-2009) is 334 mm,
with most annual precipitation falling from April through September. The
site is grass dominated and, for the purposes of our study, we focus on
the four most abundant graminoid species: \emph{Bouteloua gracilis}
(BOGR), \emph{Hesperostipa comata} (HECO), \emph{Pascopyrum smithii}
(PASM), and \emph{Poa secunda} (POSE).

From 1932 to 1945 individual plants were identified and mapped annually
in 44 1-\(\text{m}^2\) quadrats using a pantograph. The quadrats were
distributed in six pastures, each assigned a grazing treatment of light
(1.24 ha/animal unit month), moderate (0.92 ha/aum), and heavy (0.76
ha/aum) stocking rates (two pastures per treatment). In this analysis we
account for potential differences among the grazing treatments, but do
not focus on grazing\(\times\)climate interactions. The annual maps of
the quadrats were digitized and the fates of individual plants tracked
and extracted using a computer program. Daily climate data, which we
aggregated into climate variables of interest, are available for the
duration of the data collection period (1932 - 1945) from the Miles City
airport, Wiley Field, 9 km from the study site.

@@

At both levels of inference (individual and quadrat), the building
blocks of our population models are vital rate regressions. For
individual level data we fit models for survival, growth, and
recruitment of new individuals for each species. At the quadrat level we
fit a single regression model for population growth. We describe the
statistical models separately since fitting the models required
different approaches. All models contain five climate covariate that we
chose \emph{a priori}: ``water year'' precipitation at \emph{t}-1
(lagppt); fall through spring precipitation at \emph{t}-1 and \emph{t}-2
(ppt1 and ppt2, respectively) and mean spring temperature at \emph{t}-1
and \emph{t}-2 (TmeanSpr1 and TmeanSpr2, respectively), where \emph{t}
is the observation year. We also include interactions among same-year
climate covariates (e.g., ppt1 \(\times\) TmeansSpr1) and climate
\(\times\) size interactions.

We fit all models using a hierarchical Bayesian approach, which we
describe in more detail below. However, for each vital rate statistical
model we also define the likelihood model we use. For the likelihood
models, \textbf{Y} is always the relevant vector of observations (e.g.,
whether a genet survived {[}1{]} or not {[}0{]} from year \(t\) to
\(t+1\)).

@@

where \(x_{jq,t}\) is the log of species' \(j\) proportional cover in
quadrat \(q\) at time \(t\) and all other parameters are as in the
individual-level growth model (Eq. \#). The log normal likelihood
includes a truncation (T{[}0,1{]}) to ensure that predicted values do
not exceed 100\% cover.

@@

We use an environmentally stochastic IPM (Rees and Ellner 2009) that
includes the random year effects and the climate covariates from the
vital rate statistical models. But note that we can, and do for some
simulations, ignore the random year effects so that only the climate
effects can drive interannual variation. Our IPM follows the
specification of Chu and Adler (2015) where the population of species
\emph{j} is a density function \(n(u_{j},t)\) giving the density of
sized-\emph{u} genets at time \emph{t}. Genet size is on the natural log
scale, so that \(n(u_{j},t)du\) is the number of genets whose area (on
the arithmetic scale) is between \(e^{u_{j}}\) and \(e^{u_{j}+du}\). So,
the density function for any size \emph{v} at time \(t+1\) is

@@

where \(k_{j}(v_{j},u_{j},\bar{\bold{w_{j}}})\) is the population kernal
that describes all possible transitions from size \(u\) to \(v\) and
\(\bar{\bold{w_{j}}}\) is a vector of estimates of average crowding
experienced from all other species by a genet of size \(u_j\) and
species \(j\). The integral is evaluated over all possible sizes between
predefined lower (\emph{L}) and upper (\emph{U}) size limits that extend
beyond the range of observed genet sizes.

@@

which, said plainly, means we are calculating growth (\emph{G}) for
individuals that survive (\emph{S}) from time \emph{t} to \emph{t+1} and
adding in newly recruited (\emph{R}) individuals of an average sized
one-year-old genet for the focal species. Our stastical model for
recruitment (\emph{R}, described above) returns the number of new
recruit produced per quadrat. Following previous work (Adler et al.
2012, Chu and Adler 2015), we assume that fecundity increases linearly
with size
(\(R_j(v_{j},u_{j},\bar{\bold{w_{j}}}) = e^{u_j}R_j(v_{j},\bar{\bold{w_{j}}})\))
to incorporate the recruitment function in the spatially-implicit IPM.

@@

\begin{table}[ht]
\centering
\caption{Mean absolute error (MAE) and accuracy (Pearson's $\rho$) for one-step-ahead forecast from both model types. Forecasts were made without random year effects; only climate covariates could explain year-to-year variation. 90\% Distance refers to the average distance between the upper and lower 90th percentiles of the 100 predicted values for each quadrat-year combination.} 
\begin{tabular}{llrrr}
  \hline
Species & Model & MAE & 90\% Distance & Mean Obs. Cover \\ 
  \hline
BOGR & IPM & 12.18 & 38.52 & 9.43 \\ 
  BOGR & QBM & 19.66 & 56.50 & 9.26 \\ 
  HECO & IPM & 1.22 & 6.47 & 1.15 \\ 
  HECO & QBM & 12.35 & 41.11 & 1.18 \\ 
  PASM & IPM & 0.19 & 1.65 & 0.42 \\ 
  PASM & QBM & 0.55 & 7.78 & 0.42 \\ 
  POSE & IPM & 1.37 & 7.64 & 1.25 \\ 
  POSE & QBM & 1.79 & 40.59 & 1.27 \\ 
   \hline
\end{tabular}
\end{table}

The IPM MAE is significantly lower for HECO
(\(p = 3.3 \times 10^{-10}\)) and POSE (\(p = 0.00013\)). MAEs are
statisticially similar between models for BOGR and PASM.

@@

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-figure_3.pdf}
\caption{Proportional change in species' mean cover caused by a 1\%
increase in observed precipitation (+ppt), temperature (+temp), or both
(+ppt\&temp) as predicted by the individual-based IPM and the
aggregate-based QBM. Top panels show the mean predicted proportional
change in cover; lower panels show the 90\% quantiles of predicted
proportional cover change. Mean predictions and their associated
uncertainties are shown in separate panels for visual clarity.}
\end{figure}

@@

\subsection*{References}\label{references}
\addcontentsline{toc}{subsection}{References}
